Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-11-12 WSSSPE3/report/wg_appendix_best_practices.tex
small formatting/etc issues

\begin{itemize}
\item Abani Patra 
\item Sandra Gesing -- University of Notre Dame
\item Neil Chue Hong 
\item Greg Tucker 
\item Birgit Penzens 
\item Abigail Cabunoc Mayes 
\item Jeff Carver 
\item Frank L\"offler 
\item Colin Venter 
\item Lorraine Hwang 
\item Sou-Cheng Choi
\item Suresh Marru 
\item Don Middleton 
\item Daniel Katz  
\item Kyle Niemeyer 
\item Jeff Carver 
\item Dan Gunter 
\item Alexander Konovalov 
\item Tom Crick

@@

\subsection{Summary of Discussion}
Core questions that will need to be explored are in knowledge management, 
(transitions between people), reliability (reproducibility), usability and how a software tool becomes part of the core workflow of well identified users (stakeholders)
relating to tool success and hence sustainability. Ideas 
that may need to be explored include:
\begin{itemize}
\item Requirements engineering to create tools with immediate uptake;
\item When should software "die"?
\item Catering to disruptive developments in environment e.g.(new hardware, new methodology) ;
\item Dimensions of sustainability - economic, technical, environmental, 
declining interest in primary application area), social.
\item 
\end{itemize}

@@

\subsection{Key Milestones and Responsible Parties}
\item [15 Nov] Introduction and scope finished (Abani Patra/Sandra Gesing)
\item [15 Nov] Sections assigned (Abani Patra/Sandra Gesing)
\item [31 Jan] Analysing funding possibilities for survey
\item [31 Jan] First versions of section
\item [15 Feb] Distribution to WSSSPE community
\item [31 Mar] Final version of white paper
\item [30 Apr] Submission of peer-reviewed paper?

>>>>

\begin{itemize}
\item Abani Patra 
\item Sandra Gesing -- University of Notre Dame
\item Neil Chue Hong 
\item Greg Tucker 
\item Birgit Penzens 
\item Abigail Cabunoc Mayes 
\item Jeff Carver 
\item Frank L\"{o}ffler 
\item Colin Venter 
\item Lorraine Hwang 
\item Sou-Cheng Choi
\item Suresh Marru 
\item Don Middleton 
\item Daniel Katz  
\item Kyle Niemeyer 
\item Jeff Carver 
\item Dan Gunter 
\item Alexander Konovalov 
\item Tom Crick

@@

\subsection{Summary of Discussion}
Core questions that will need to be explored are in knowledge management, 
(transitions between people), reliability (reproducibility), usability and how a software tool becomes part of the core workflow of well identified users (stakeholders)
relating to tool success and hence sustainability. Ideas 
that may need to be explored include:
\begin{itemize}
\item Requirements engineering to create tools with immediate uptake;
\item When should software ``die''?
\item Catering to disruptive developments in environment e.g.(new hardware, new methodology) ;
\item Dimensions of sustainability - economic, technical, environmental, 
declining interest in primary application area), social.
\end{itemize}

@@

\subsection{Key Milestones and Responsible Parties}
\begin{itemize}
\item 15 Nov: Introduction and scope finished (Abani Patra/Sandra Gesing)
\item 15 Nov: Sections assigned (Abani Patra/Sandra Gesing)
\item 31 Jan: Analyzing funding possibilities for survey
\item 31 Jan: First versions of section
\item 15 Feb: Distribution to WSSSPE community
\item 31 Mar: Final version of white paper
\item 30 Apr: Submission of peer-reviewed paper?
\end{itemize}

### 2015-11-12 WSSSPE3/report/wg_main_best_practices.tex
small formatting/etc issues

\subsubsection{Why it is important}
Reviewing multiple past articles and talks at different meetings like WSSSPEx\cite{WSSSPE papers, heroux paper, ...} and 
analyzing and promoting sustainable scientific software makes it clear that there 
are several common and recurring ideas that underpin success. However, outside of a small community, this knowledge is not widely
shared. This is  especially true for the large community of scientists who generate most of the software used by scientists but are not primarily software developers. In this
scenario a clear and precise exposition of these best practices collected from many sources and open collaboration among all in the community
in a single source (e.g. a journal paper, tutorial etc.)  that can be widely disseminated is necessary and likely to be very valuable.

@@

The creation of such a ``best practices" document will build upon the range of activities and topics discussed at WSSSPE3 and associated prior meeting. We will attempt to
distill the emerging body of knowledge into this document. The large number of  articles from the NSF funded S2I2 projects (SSE and SSI) ``lightning talks", ``white papers"  and reports from different workshops point have created a large if somewhat diffuse source for this report.

@@

Core questions that will need to be explored are in knowledge management, 
(transitions between people), reliability (reproducibility), usability and how a software tool becomes part of the core workflow of well identified users (stakeholders)
relating to tool success and hence sustainability. Ideas 
that may need to be explored include:
\begin{itemize}
\item Requirements engineering to create tools with immediate uptake;
\item When should software "die"?
\item Catering to disruptive developments in environment e.g.(new hardware, new methodology) ;
\item Dimensions of sustainability - economic, technical, environmental, 
declining interest in primary application area), social.
\item 
\end{itemize}

@@

@@

\noindent{\bf Timeline:}
\begin{itemize}
%
\item [15 Nov] Introduction and scope finished
\item [15 Nov] Sections assigned
\item [31 Jan] Analysing funding possibilities for survey
\item [31 Jan] First versions of section
\item [15 Feb] Distribution to WSSSPE community
\item [31 Mar] Final version of white paper
\item [30 Apr] Submission of peer-reviewed paper?
\end{itemize}

@@

\subsubsection{Landing Page}
The landing page with instructions, timeline and the white paper is here: https://drive.google.com/drive/folders/0B7KZv1TRi06fbnFkZjQ0ZEJKckk
Discussions can be also continued in https://github.com/WSSSPE/meetings/issues/42

>>>>

%\subsubsection{Why it is important}
Reviewing multiple past articles and talks at different meetings like WSSSPEx\cite{WSSSPE papers, heroux paper, ...} and 
analyzing and promoting sustainable scientific software makes it clear that there 
are several common and recurring ideas that underpin success in developing sustainable software. However, outside of a small community, this knowledge is not widely
shared. This is  especially true for the large community of scientists who generate most of the software used by scientists but are not primarily software developers. In this
scenario a clear and precise exposition of these best practices collected from many sources and open collaboration among all in the community
in a single source (e.g. a journal paper, tutorial etc.)  that can be widely disseminated is necessary and likely to be very valuable.

@@

The creation of such a ``best practices'' document will build upon the range of activities and topics discussed at WSSSPE3 and associated prior meeting. We will attempt to
distill the emerging body of knowledge into this document. The large number of  articles from the NSF funded S2I2 projects (SSE and SSI) ``lightning talks'', ``white papers''  and reports from different workshops point have created a large if somewhat diffuse source for this report.

@@

Core questions that will need to be explored are in knowledge management, 
(transitions between people), reliability (reproducibility), usability and how a software tool becomes part of the core workflow of well identified users (stakeholders)
relating to tool success and hence sustainability. Ideas 
that may need to be explored include:
\begin{itemize}
\item Requirements engineering to create tools with immediate uptake;
\item When should software ``die''?
\item Catering to disruptive developments in environment e.g.(new hardware, new methodology) ;
\item Dimensions of sustainability - economic, technical, environmental, 
declining interest in primary application area), social.
\end{itemize}

@@

For additional information about the discussion, see Appendix~\ref{sec:appendix_best_practices}.

@@

\noindent{\bf Timeline:}
\begin{itemize}
%
\item 15 Nov: Introduction and scope finished
\item 15 Nov: Sections assigned
\item 31 Jan: Analyzing funding possibilities for survey
\item 31 Jan: First versions of section
\item 15 Feb: Distribution to WSSSPE community
\item 31 Mar: Final version of white paper
\item 30 Apr: Submission of peer-reviewed paper?
\end{itemize}

@@

\subsubsection{Landing Page}
The landing page with instructions, timeline and the white paper is here: \url{https://drive.google.com/drive/folders/0B7KZv1TRi06fbnFkZjQ0ZEJKckk}
Discussions can be also continued in \url{https://github.com/WSSSPE/meetings/issues/42}
